
\chapter{Benefits and Drawbacks}

There are many benefits to using cryptocurrencies such as Bitcoin,
some apply now other may come to be if and when cryptocurrencies become
more adopted in more conventional places. Some key benefits include:
\begin{itemize}
\item a self-regulating system, no unexpected interventions
\item no need for middlemen, such as banks
\item virtually instant transactions
\end{itemize}

\section*{A Self-Regulating System}

The Bitcoin system is self-regulating, meaning that no government
or institution controls it. For the user that means that their holdings
cannot be used as a policy instrument. Conventional fiat currencies
are owned by a government, which means that they can be used to e.g.
stimulate economic growth by printing extra money, which lessens the
value of individuals holdings.


\section*{No middlemen}

Secondly, users control their own holdings on their own computer,
which eliminates the need for a bank. This brings the advantage that
users are not dependent on open hours, waiting lines, or e-banking
websites, it also means no fees. Additionally a user's holdings cannot
be wiped out by his or her bank bankrupting after speculative investing
or something of that sort.


\section*{Virtually Instant Transactions}

Thirdly, transactions are as easy, instant, and costless as sending
an email. Bitcoin exists online digitally, which means that there
is no physical counterpart that needs to be moved, such as with gold,
or paper money. Additionally, it is equally valid across countries,
which means that there is no need to exchange it. 


